One Model May Not Fit All: Subgroup Detection Using MOB

### 2023-10-04 _MOB_paper.Rnw
Comment on glmertree example

With the first argument, we specified the model \verb|formula|, which has three parts separated by vertical bars: The left part (\verb|PPVT ~ Program*Age|) specifies the response variable, followed by a tilde (\verb|~|) and the fixed-effects predictors of relevance. The middle part (\texttt{1|ChildID}) specified the random effects. The right part (\verb|AFTQ + Race + Income + Mom_edu_yrs + Mom_height|) specified the partitioning variables: covariates that may possibly affect the values of the fixed-effects parameters.

With the second argument, we specified the dataset which contain the variables. Because we are dealing with repeated measurements on the same children, we additionally specified that the parameter stability tests should be performed on the child level using the \verb|cluster| argument. Using the default observation-level parameter stability tests may artificially inflate power. Finally, because we want to retain large enough subgroups, we specified that the minimum number of observations in a terminal node should be 250.

@@

To aid interpretation of the coefficients in Figure~\ref{fig:lmm_tree}, Table~\ref{tab:predictions} provides predicted PPVT scores for each of the groups and programs at ages 6 \todo{x axis for age starts at 2?} and 18. All nodes show a modest benefit of Head Start participation at age 6, about 3-4 points on the PPVT. This benefit remains the same over time for the group with lower mother's AFTQ scores. The benefit increases over time for White children with higher mother's AFTQ scores. Strikingly,\todo{possibly less extreme word like: Interestingly,} the benefit decreases over time for Black and Hispanic children with higher mother's AFTQ scores. These results correspond to the conclusions of \cite{Demi09}. \todo[inline]{Generally the effects are not large but the example is very clear and I like that it is an intervention effect. I suggest to state early before the figure that the effects are not large in this example but we are interested in the group differences; also try to make figure bigger and wider to make slopes more visible}

@@

\todo[inline]{I am not sure this dataset is the best GLMM tree illustration, because visualized differences between subgroups are very small and perhaps the fixed-effect part is already too complex. Alternative: Do not focus on effect of Head Start, but use as partitioning variable. Then can also use data from more families (HS or none not needed), partition using child-level characteristics, and  model specification becomes simpler. -- response Caro: I like it, see comment above: try to prime that effects are not large}

>>>>

With the first argument, we specified the model \verb|formula|, which has three parts separated by vertical bars: The left part (\verb|PPVT ~ Program*Age|) specifies the response variable, followed by a tilde (\verb|~|) and the fixed-effects predictors of relevance. The middle part (\texttt{1|ChildID}) specified the random effects\todo{(Mirka) Maybe add information that this is for each child, average DV across timepoints}. The right part (\verb|AFTQ + Race + Income + Mom_edu_yrs + Mom_height|) specified the partitioning variables: covariates that may possibly affect the values of the fixed-effects parameters.

\todo[inline]{I (Mirka) am not sure what exactly the cluster argument does. Maybe Marjolein you have more information from your work with Achim. But my knowledge/intuition is that it has something to do with clustered covariances, so maybe it accounts for the sibling structure, but I don't think it accounts for the level of the predictor variables.}
With the second argument, we specified the dataset which contain the variables. Because we are dealing with repeated measurements on the same children, we additionally specified that the parameter stability tests should be performed on the child level using the \verb|cluster| argument. Using the default observation-level parameter stability tests may artificially inflate power\todo{Mirka: I'd rather say it has an inflated type-I error rate.}. Finally, because we want to retain large enough subgroups, we specified that the minimum number of observations in a terminal node should be 250.

@@

To aid interpretation of the coefficients in Figure~\ref{fig:lmm_tree}, Table~\ref{tab:predictions} provides predicted PPVT scores for each of the groups and programs at ages 6 \todo{x axis for age starts at 2?} and 18. All nodes show a modest benefit of Head Start participation at age 6, about 3-4 points on the PPVT. This benefit remains the same over time for the group with lower mother's AFTQ scores. The benefit increases over time for White children with higher mother's AFTQ scores. Strikingly,\todo{possibly less extreme word like: Interestingly,} the benefit decreases over time for Black and Hispanic children with higher mother's AFTQ scores. These results correspond to the conclusions of \cite{Demi09}. \todo[inline]{Generally the effects are not large but the example is very clear and I like that it is an intervention effect. I suggest to state early before the figure that the effects are not large in this example but we are interested in the group differences; also try to make figure bigger and wider to make slopes more visible; (Mirka): I think the main effect of the intervention and the interaction effect is not too small, but it is hard to see because the age effect is so pronounced in this example. Would it be an option to only use the intervention effect in the end nodes? }

@@

\todo[inline]{I am not sure this dataset is the best GLMM tree illustration, because visualized differences between subgroups are very small and perhaps the fixed-effect part is already too complex. Alternative: Do not focus on effect of Head Start, but use as partitioning variable. Then can also use data from more families (HS or none not needed), partition using child-level characteristics, and  model specification becomes simpler. -- response Caro: I like it, see comment above: try to prime that effects are not large -- response Mirka: I think that's okay. Only if you had data at hand that is not longitudinal (rather students in classes/schools), with covariates on Level-1, I would opt for that because it doesn't have the Level-2 splitting issue. }
